---
title: "Lab 9 - Matrix Modeling"
author: "Linda Dominguez"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
# This chunk installs weplot and helps with pdf formatting (and is excluded from knitting)
source("setup/setup.R")
source("setup/weplot.R")
load("Heliconia.RData")
```


## Part 1 – A perennial plant with a simple life cycle

### (1a) Define your projection matrix

```{r 1a}
# A<- c(0, 2.4, 7.2,.2, .21, .06,0, .11, .42)
#turn into matrix w 3 rows and cols 
A <- matrix(c(0, 2.4, 7.2, .2, .21, .06, 0, .11, .42), nrow = 3, byrow = TRUE)

#add names to rows & col 
rownames(A)<- c("Sdlg", 'Sm', 'Lg')
colnames(A)<- c("Sdlg", 'Sm', 'Lg')

A

```


### (1b) Make a model that projects this population 10 years into the future

```{r 1b}
#represent N as matrix w rows = stage class, cols = time

time<- 10 #bc we want 10 yr into future => same as Time
rowCol <- c(3, time+1) #same as Row.col
N<- array(0, rowCol) #makes 2D array of zeros
rownames(N)<- c("Sdlg", 'Sm', 'Lg') #names stafes
colnames(N)<- c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10)

#insert staring pop 
N[,1]<- c(290, 55, 12)
N
# N[6] <- 16 #sets individual at 6th place  to be 16 e.g. => just exploring to learn R better 


populationProjection<- function(time, startingPop){
  rowCol <- c(3, time+1) #same as Row.col
  N<- array(0, rowCol) #makes 2D array of zeros
  rownames(N)<- c("Sdlg", 'Sm', 'Lg') #names stafes
  colnames(N)<- c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10)

  #insert staring pop 
  N[,1]<- c(startingPop)
  for (t in 1:time){
  N[,t+1]<- A%*%N[,t]}

  return (N)
}

startPop<- c(290, 55, 12)
N <- populationProjection(time, startPop)
N


```


### (1c) What is the projected population size vector after 10 years?
The projected population sizes are: 
  Sdlg        Sm        Lg
217.73375  56.75444  10.97858

```{r 1c}
projectedN10yr<- N[,11] #11 bc/ it starts at 1 not 0 
projectedN10yr
#   Sdlg        Sm        Lg
# 217.73375  56.75444  10.97858


```


### (1d) Make a figure showing the numbers of each stage class through time

```{r 1d}
stagesThroughTimePlot<- weplot.MM(N) #weplot.MM is used to plot matrix model 
stagesThroughTimePlot


```


### (1e) What is the long-term Lambda value for this population based on your one year of data?

```{r 1e}
totalN <- apply(N, 2, sum) #applies sum funct between cols 
longTermLambda<- totalN[time+1]/totalN[time]
longTermLambda

#second way of finding lambda 
longTermLambda2<- as.numeric(eigen(A)$values[1])
longTermLambda2

getLambda<- function(x){
  return(as.numeric(eigen(x)$values[1]))
  }
longTermLambda2_1<-getLambda(A)
longTermLambda2_1


```


### (1f) Examine two scenarios where your population begins with very different starting population size vectors
* Both Scenario 1 and Scenario 2 have fairly consistent populations.
* Scenario 2 has population magnitude in the thousands, ranging from whar appears to be (0, 36000).
* Scenario 1 has population magnitude in the thousands too, but ranging from whar appears to be (0, 5000).
* Scenario 1 which has a larger starting Seedling population ends up having a population about 1/10 of the magnitude of Scenario 2 which started with a bigger Large Reproductive population.
* Scenario 2 shows that a bigger Large Reproductive population leads to a larger boom i population.

```{r 1f}

#For each scenario, be sure to include text that *interprets* why you think the population sizes behave the way that they do.
scenario1 <- populationProjection(10, c(5000, 55, 12))
scenario1
scenario2 <- populationProjection(10, c(290, 55, 5000))
scenario2

#plotting 
scenario1Plot<- weplot.MM(scenario1, title='Scenario1')
scenario2Plot <- weplot.MM(scenario2, title='Scenario 2')

```


## Part 2 – An overlooked part of the lifecycle?

### (2a) Modify your projection matrix (A) from Part 1 to incorporate a seed bank

```{r 2a}


populationProjection2 <- function(time, startingPop, aValues){

  A <- matrix(c(aValues), nrow = 4, byrow = TRUE)

  #add names to rows & col 
  rownames(A)<- c("Dormant","Sdlg", 'Sm', 'Lg')
  colnames(A)<- c("Dormant","Sdlg", 'Sm', 'Lg')

  rowCol <- c(4, time+1) #same as Row.col
  N<- array(0, rowCol) #makes 2D array of zeros
  rownames(N)<- c("Dormant", "Sdlg", 'Sm', 'Lg') #names stafes
  colnames(N)<- c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10)

  #insert staring pop 
  N[,1]<- c(startingPop)
  for (t in 1:time){
  N[,t+1]<- A%*%N[,t]}
  print(c("lambda ", getLambda(A)))
  return (N)
}


Avalues<- c(.22, 0, 9.1, 18.5, .1, 0, 2.4, 7.2, 0, .2, .21, .06, 0, 0, .11, .42)
startingPop <- c(1090, 290, 55, 12)
seedBankPop <- populationProjection2(10, startingPop, Avalues)
seedBankPop


```


### (2b) After including the presence of a seed bank, what is the revised Lambda value for the population?
The revised lambda is: 1.10707276020224.
```{r 2b}



```


### (2c) Create a figure of the projected numbers for all stage classes over 10 years

```{r 2c}
dormantPlot <- weplot.MM(seedBankPop, title='Projected Population Over 10 years Including Dormant Seeds')



```


## Part 3 – Habitat fragmentation in the tropics

### (3a) Use the three different projection matrices (A.Sm, A.Med, A.Con) to determine the estimated Lambda values in the different sized forest habitats.
Small Plot Lambda:0.989879, Medium Plot Lambda:1.002689, Continuous Plot Lambda:1.029313
```{r 3a}
load('Heliconia.RData')
# ls()
#finding lambdas 
smallPlotLam <- getLambda(A.Sm)
smallPlotLam
medPlotLam <- getLambda(A.Med)
medPlotLam
continuousPlotLam<- getLambda(A.Con)
continuousPlotLam
# > smallPlotLam
# [1] 0.989879
# > medPlotLam
# [1] 1.002689
# > continuousPlotLam
# [1] 1.029313
```


### (3b) Create a population model for the smallest habitat fragment (1 ha) that projects the population out 10 years

```{r 3b}


populationProjection3 <- function(time, aValues, rowNames){
  A <- aValues

  rowCol <- c(6, time+1) #same as Row.col
  N<- array(0, rowCol) #makes 2D array of zeros
  rownames(N)<- rowNames #get names from N.Sm
  colnames(N)<- c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10)


  #insert staring pop 
  N[,1]<- rowNames
  for (t in 1:time){
  N[,t+1]<- A%*%N[,t]}
  print(c("lambda ", getLambda(A)))
  return (N)
}
A<- A.Sm #matrix data from smallest habitat fragment (A.Sm)
rowNames <- rownames(N.Sm)
HeliconaModel <- populationProjection3(10, A, rowNames)
HeliconaModel


```


### (3c) Make a plot showing the projected numbers for all stage classes out for 10 years

```{r 3c}

HeliconaModelPlot <- weplot.MM(HeliconaModel, title = "Heliconia Acuminata 1 ha Model")



```


### 3d) Let’s incorporate more detail into the values that represent reproduction in Heliconia (i.e. the first row of the projection matrix)

```{r 3d}




```


### (3e) Assuming that total seed production remains the same, what germination probability would give Lambda = 1.0 in the forest fragments?

```{r 3e}




```


### (3f) Modify your model so that you can incorporate the (external) addition of seeds into the plot each year.  How many seeds would you have to add each year to have a stable population after 10 years?

```{r 3f}




```


### (3g) How many seeds must be dispersing into the plots in order to produce an observed Lambda value of 1.06 one year later?

```{r 3g}




```
